---
title: "Week7"
author: "Samantha Gouveia"
date: "`r Sys.Date()`"
output: html_document
---
```{r}
library(knitr)
   library(car)
   library(dplyr)
   library(kableExtra)
   library(pander)
```

#Data Introduction 

The data we are using is from the U.S. Small Business Administration. It provides historical data from 1987 until 2014. 

There are 27 variables and 899,164 observations each representing a loan.


```{r}

loan01 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational01.csv", header = TRUE)[, -1]
loan02 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational02.csv", header = TRUE)[, -1]
loan03 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational03.csv", header = TRUE)[, -1]
loan04 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational04.csv", header = TRUE)[, -1]
loan05 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational05.csv", header = TRUE)[, -1]
loan06 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational06.csv", header = TRUE)[, -1]
loan07 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational07.csv", header = TRUE)[, -1]
loan08 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational08.csv", header = TRUE)[, -1]
loan09 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational09.csv", header = TRUE)[, -1]
loan = rbind(loan01, loan02, loan03, loan04, loan05, loan06, loan07, loan08, loan09)
```

# Remove missing values

We begin by removing any loans that are missing an MIS Status.

```{r Remove Missing Values}

###1 Remove missing MIS_Status values 
nloan <- na.omit(loan)
sum2 <- sum(is.na(nloan$MIS_Status))

kable(sum2, format = "markdown", col.names = c("Number of Missing MIS_Status Values"))
```

When looking at the data we find no loans are missing an MIS Status.

# Stratifing the Population

We stratify the population based on the region the loan is in. There are 7 regions we will look at
- New England 
- Mid-Atlantic
- South
- Great Lakes
- Midwest
- Pacific Coast
- Non Continental
- Rocky Mountains

By stratifying the data into regions we have created 7 sub populations. This will allow us to see a more accurate result due to there being smaller sample sizes.

```{r}

#Remove observations w/o state classification
nloan$State[nloan$State == ""] <- NA
nloan <- na.omit(nloan)


cnvrt <- c(
   "ME" = "New England", "NH" = "New England", "VT" = "New England", "MA" = "New England", "RI" = "New England", "CT" = "New England",
   "NJ" = "Mid-Atlantic", "DE" = "Mid-Atlantic", "MD" = "Mid-Atlantic", "DC" = "Mid-Atlantic",  
   "WV" = "South", "VA" = "South", "NC" = "South", "SC" = "South", "GA" = "South", "FL" = "South", "KY" = "South", "TN" = "South", "AL" = "South", "MS" = "South", "AR" = "South", "LA" = "South",
   "MN" = "Great Lakes", "WI" = "Great Lakes", "MI" = "Great Lakes", "OH" = "Great Lakes ", "IN" = "Great Lakes", "IL" = "Great Lakes", "IA" = "Great Lakes","NY" = "Great Lakes",  "PA" = "Great Lakes", "MO" = "Midwest",
   "ND" = "Midwest", "SD" = "Midwest", "NE" = "Midwest", "KS" = "Midwest ", "OK" = "Midwest",
   "MT" = "Rocky Mountain", "WY" = "Rocky Mountain", "CO" = "Rocky Mountain", "UT" = "Rocky Mountain", "ID" = "Rocky Mountain",
   "NV" = "Southwest", "AZ" = "Southwest", "NM" = "Southwest", "TX" = "Southwest",
   "WA" = "Pacific Coast", "OR" = "Pacific Coast", "CA" = "Pacific Coast",
   "AK" = "Non Continental", "HI" = "Non Continental"
  
)

nloan$Region <- cnvrt[nloan$State]

cregion <- nloan %>%
  count(Region) %>%
  rename(Count = n)

kable_region <- kable(cregion, format = "markdown")

kable_region
```

# Defining Clusters

We will divide each sub population into clusters bases on there zip code. The different loans within each region will be clustered as it allows a closer look at loans in similar areas.

zip code
```{r}
nloan <- nloan[!is.na(nloan$Zip), ]


num_clusters <- 7

kmeans_result <- kmeans(nloan$Zip, centers = num_clusters)

cluster_centers <- kmeans_result$centers

nloan$Cluster <- kmeans_result$cluster

print(cluster_centers)

print(head(nloan))



```

#Simple Random Sampling 

To take a look at the study population we take random samples. This allows us to observe the data without any bias. 

```{r}

set.seed(123)  
random_sample <- sample_n(nloan, size = 100)  
```

# Systematic Sample

A systematic samples allows for the random sample to be fixed with a fixed interval of 50. 

By using the systematic sample there will be no influence over which loans are chosen allowing for the study to be unbiased.
```{r}

systematic_sample <- nloan[seq(1, nrow(nloan), by = 50), ]
```

# Stratified Sample

A stratified sample will be done using the 7 regions we created earlier in the study.

A sample size of 100 will be taken from each region and used in the final study population. This allows for there to be little bias as there is an equal amount of samples from each regions.
```{r}

stratified_sample <- nloan %>%
  group_by(Region) %>%
  sample_n(size = 100)  # Adjust the sample size as needed
```

# Cluster by Zip Code

We cluster by zip code using the entire study population. We will create clusters and select 7 of them to be used as the final study population.

When clustering by zip code it is important for the researchers to know we have a chance of getting no data from smaller areas as they may be picked less often.

```{r}

clusters <- unique(nloan$Cluster)

selected_clusters <- sample(clusters, size = 7)

cluster_sample <- nloan %>%
  filter(Cluster %in% selected_clusters)

cat("Random Sample:", nrow(random_sample), "observations\n")
cat("Systematic Sample:", nrow(systematic_sample), "observations\n")
cat("Stratified Sample:", nrow(stratified_sample), "observations\n")
cat("Cluster Sample:", nrow(cluster_sample), "observations\n")
```

# Conclusion

Throughout the study we have looked at the data provided by the U.S. Small Business Administration. We have looked at 4 different sampling methods to analysis the loans. When completing the study we do not find that any method is better than another.
