---
title: "Exploratory Data Analysis and Data Preprocessing"
author: "Samantha Gouveia"
date: "03/08/2024"
output:
 html_document:
  toc: yes
  toc_depth: 4
  toc_float: yes
  fig_width: 6
  fig_caption: yes
  number_sections: yes
  theme: readable
  editor_options:
  chunk_output_type: console
---




```{=html}

<style type="text/css">

/* Cascading Style Sheets (CSS) is a stylesheet language used to describe the presentation of a document written in HTML or XML. it is a simple mechanism for adding style (e.g., fonts, colors, spacing) to Web documents. */

h1.title {  /* Title - font specifications of the report title */
  font-size: 24px;
  color: DarkRed;
  text-align: center;
  font-family: "Gill Sans", sans-serif;
}
h4.author { /* Header 4 - font specifications for authors  */
  font-size: 20px;
  font-family: system-ui;
  color: DarkRed;
  text-align: center;
}
h4.date { /* Header 4 - font specifications for the date  */
  font-size: 18px;
  font-family: system-ui;
  color: DarkBlue;
text-align: center;
}
h1 { /* Header 1 - font specifications for level 1 section title  */
    font-size: 22px;
    font-family: system-ui;
    color: navy;
    text-align: left;
}
h2 { /* Header 2 - font specifications for level 2 section title */
    font-size: 20px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h3 { /* Header 3 - font specifications of level 3 section title  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h4 { /* Header 4 - font specifications of level 4 section title  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: left;
}

body { background-color:white; }

.highlightme { background-color:yellow; }

p { background-color:white; }

</style>
```

```{r setup, include=FALSE}

library(knitr)
   library(car)
   library(dplyr)
   library(kableExtra)
   library(pander)
knitr::opts_chunk$set(echo = TRUE,  
                      warning = FALSE, 
                      result = TRUE,
                      message = FALSE,
                      comment = NA
                      )   
```



#Data Introduction 

The data we are using is from the U.S. Small Business Administration. It provides historical data from 1987 until 2014. 

There are 27 variables and 899,164 observations each representing a loan.

```{r}
loan01 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational01.csv", header = TRUE)[, -1]
loan02 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational02.csv", header = TRUE)[, -1]
loan03 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational03.csv", header = TRUE)[, -1]
loan04 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational04.csv", header = TRUE)[, -1]
loan05 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational05.csv", header = TRUE)[, -1]
loan06 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational06.csv", header = TRUE)[, -1]
loan07 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational07.csv", header = TRUE)[, -1]
loan08 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational08.csv", header = TRUE)[, -1]
loan09 = read.csv("https://pengdsci.github.io/datasets/SBAloan/w06-SBAnational09.csv", header = TRUE)[, -1]
loan = rbind(loan01, loan02, loan03, loan04, loan05, loan06, loan07, loan08, loan09)
# dim(bankLoan)
#names(bankLoan)
```

# Remove missing values

We begin by removing any loans that are missing an MIS Status.

```{r Remove Missing Values}

###1 Remove missing MIS_Status values 
nloan <- na.omit(loan)
sum2 <- sum(is.na(nloan$MIS_Status))

kable(sum2, format = "markdown", col.names = c("Number of Missing MIS_Status Values"))
```

We find there are 0 observations missing an MIS Status value.

# Formatting variables 

We look at 5 variables. DisbursementGross, BalanceGross, ChgOffPrinGr, GrAppv, and SBA_Appve and change them to be regular numeric variables. 
```{r}
currency_vars <- c('DisbursementGross', 'BalanceGross', 'ChgOffPrinGr', 'GrAppv', 'SBA_Appv')

for (var in currency_vars) {
  loan[, var] <- as.numeric(gsub("[\\$,]", "", loan[, var]))
}

```

# Stratifing the Population

We stratify the population based on the region the loan is in. There are 7 regions we will look at
- New England 
- Mid-Atlantic
- South
- Great Lakes
- Midwest
- Pacific Coast
- Non Continental
- Rocky Mountains

By stratifying the data into regions we have created 7 sub populations. This will allow us to see a more accurate result due to there being smaller sample sizes.

```{r}

#Remove observations w/o state classification
nloan$State[nloan$State == ""] <- NA
nloan <- na.omit(nloan)


cnvrt <- c(
   "ME" = "New England", "NH" = "New England", "VT" = "New England", "MA" = "New England", "RI" = "New England", "CT" = "New England",
   "NJ" = "Mid-Atlantic", "DE" = "Mid-Atlantic", "MD" = "Mid-Atlantic", "DC" = "Mid-Atlantic",  
   "WV" = "South", "VA" = "South", "NC" = "South", "SC" = "South", "GA" = "South", "FL" = "South", "KY" = "South", "TN" = "South", "AL" = "South", "MS" = "South", "AR" = "South", "LA" = "South",
   "MN" = "Great Lakes", "WI" = "Great Lakes", "MI" = "Great Lakes", "OH" = "Great Lakes ", "IN" = "Great Lakes", "IL" = "Great Lakes", "IA" = "Great Lakes","NY" = "Great Lakes",  "PA" = "Great Lakes", "MO" = "Midwest",
   "ND" = "Midwest", "SD" = "Midwest", "NE" = "Midwest", "KS" = "Midwest ", "OK" = "Midwest",
   "MT" = "Rocky Mountain", "WY" = "Rocky Mountain", "CO" = "Rocky Mountain", "UT" = "Rocky Mountain", "ID" = "Rocky Mountain",
   "NV" = "Southwest", "AZ" = "Southwest", "NM" = "Southwest", "TX" = "Southwest",
   "WA" = "Pacific Coast", "OR" = "Pacific Coast", "CA" = "Pacific Coast",
   "AK" = "Non Continental", "HI" = "Non Continental"
  
)

nloan$Region <- cnvrt[nloan$State]

cregion <- nloan %>%
  count(Region) %>%
  rename(Count = n)

kable_region <- kable(cregion, format = "markdown")

kable_region
```

# Default rate

We now find the default rates of the regions by account charged.
```{r}

default_rate <-  nloan %>%
  group_by(Region) %>%
  summarize(Count = n(),
            CHGOFF_count = sum(MIS_Status == "CHGOFF"),
            Default_Rate = (CHGOFF_count / Count) * 100
            )

kable_rdefault <- kable(default_rate, format = "markdown", col.names = c("Region", "Count", "Number Defaulted", "Default Rate"))

kable_rdefault
```


# Density Curved

```{r}
# Ensure 'SBA_Appv' is numeric and 'Region' is a factor
nloan$SBA_Appv <- as.numeric(nloan$SBA_Appv)
nloan$Region <- as.factor(nloan$Region)

# Define colors for the density curves
colors <- rainbow(length(levels(nloan$Region)))

# Initialize the plot
plot(NULL, xlim = c(0, 1000000), ylim = c(0, 1e-06), 
     main = "Density of SBA_Appv Across Regions",
     xlab = "SBA_Appv", ylab = "Density")

# Loop through each region and plot density curves
for (i in seq_along(levels(nloan$Region))) {
  region <- levels(nloan$Region)[i]
  
  # Remove NA values for the specific subset
  subset_data <- nloan$SBA_Appv[nloan$Region == region]
  subset_data <- subset_data[!is.na(subset_data)]
  
  # Plot the density if there are enough data points
  if (length(subset_data) >= 2) {
    lines(density(subset_data), col = colors[i], lwd = 2)
  } else {
    warning(paste("Not enough data points to plot density for", region))
  }
}

# Add a legend
legend("topright", legend = levels(nloan$Region), col = colors, lwd = 2)
```

Looking at the density curves we see the distribution of the different regions in relation to the SBA_Appv. We can see the spread of our discretized bins as well.

#Summary 

In conclusion we looked at data in relation to MIS Status and completed data analysis and processing.
We looked at the regions of the US in relation to the GrAppv. Through this we looked at the impact on each region based on loans.


