---
title: "Private College Acceptance"
subtitle: "Multiple linear Regression"  
author: 
  - "Samantha Gouveia"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
library(psych)
library(mlbench) 
s <- read.csv("https://raw.githubusercontent.com/sgouveia23/STA321/main/projecct2/College_Data.csv", header = TRUE)
cc <- s[,-c(8,9,11,12,13,14,15,16,17,18)]
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF"
)
```


# Data Introduction

The case study focuses on 777 private and public colleges with 8 predictors

Response variable: Private

Predictors: 
>Application

>Acceptance

>Enrollment

>Top 10 percent

>Top 25 percent

>Grad rate 

---

## Agenda

The objective of the case study is to identify the criteria needed to get into a private college.

Report aims to explore correlation between variables and explore which model will preform the best.

> We will use 

>a pairwise scatterplot

> a table

> a scatterplot


---
### Pairwise scatterplot 
```{r pairwise scatterplot}
pairs.panels(cc,method="pearson",hist.col="plum",density=TRUE,ellipses=TRUE )
```
---
```{r}
tibble::as_tibble(cc)
```
---
## Plots

```{r}
library(ggplot2)
(g <- ggplot(cc) + aes(Accept, Enroll, color = Apps)+geom_point())
```
---
